---
title: "Emirates Cup Final Analysis"
author: "Steven Martinez"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r Loading Libraries, include=FALSE}
load_libraries <- function(packages) {
  # Check for missing packages
  missing_packages <- packages[!(packages %in% installed.packages()[, "Package"])]
  
  # install missing packages
  if(length(missing_packages) > 0) {
    install.packages(missing_packages)
  }
  
  # Load all packages
  lapply(packages, library, character.only = TRUE)
  
  cat("All packages are loaded succesfully.\n")
}


# Loading necessary libraries
load_libraries(c("tidyverse", "lubridate", "stats", "ggplot2", "corrplot", "stringr", "stringi", "class",
                 "tidymodels", "modeldata", "themis", "vip", "baguette", "janitor", "rvest",
                 "yardstick", "gsheet", "caret", "randomForest", "here", "tibble", "dplyr", "ISAR",
                 "tidyr", "mgcv", "teamcolors", "baseballr", "Lahman", "remotes", "ggcorrplot", "broom",
                 "readr", "glmnet", "xgboost", "Matrix", "Metrics", "reshape2", "DMwR2", "smotefamily",
                 "highcharter", "jpeg", "grid", "gridExtra", "RCurl", "gt"))

# Load only the necessary functions from 'car'
library(car, exclude = "select")

# Turning off warning messages
options(warnings = 0)
```

```{r Loading Dataset, include=FALSE}
# Loading Dataset
df <- read_csv("data/data_mil_okc_121724_shots.csv")
```

```{r Data Cleaning, include=FALSE}
# Looking at the structure of the dataset
str(df)

# Changing some variables that are in character form into numeric
df <- df %>%
  mutate(across(c(loc_x_passer, loc_y_passer, rimDepth, rimLeftRight, arcAngle), as.numeric))


# Changing binary variables to factors
df <- df %>%
  mutate(across(where(~ is.numeric(.) && all(na.omit(.) %in% c(0, 1))),
                as.factor))

# Creating a column for 2-point makes and attempts
df <- df %>%
  mutate(
    fga2 = ifelse(fga == 1 & fga3 == 0, 1, 0),
    fg2 = ifelse(fg == 1 & fg3 == 0, 1, 0)
    )
```

```{r Half court image for shooting chart, include=FALSE}
# half court image
courtImg.URL <- "https://thedatagame.com.au/wp-content/uploads/2016/03/nba_court.jpg"
court <- rasterGrob(
  readJPEG(getURLContent(courtImg.URL)),
  width=unit(1,"npc"),
  height=unit(1,"npc")
  )
```


```{r Three pointers made by each team Shot Chart, include=FALSE}
# Shot chart for 3 pointers made in the game by each team
df %>% 
  filter(fg3 == 1) %>% 
  ggplot() +
  
  # Add court background
  annotation_custom(court, xmin = -25, xmax = 25, ymin = 0, ymax = 47) +
  
  # Shooter positions
  geom_point(aes(x = loc_x, y = loc_y), color = "blue", size = 2, alpha = 0.6) +
  
  # Passer positions
  geom_point(aes(x = loc_x_passer, y = loc_y_passer), color = "green", size = 2, alpha = 0.6) +
  
  # Pass lines
  geom_segment(aes(x = loc_x_passer, y = loc_y_passer, 
                   xend = loc_x, yend = loc_y),
               arrow = arrow(length = unit(0.12, "inches")),
               color = "gray30", alpha = 0.4) +
  
  # Facet by team
  facet_wrap(~ team_nba_off) +
  
  coord_fixed() +
  labs(title = "Three Point FGs Made Shot Chart with Passes by Team",
       subtitle = "Shooter (blue), Passer (green), Pass (gray)",
       x = "Court X", y = "Court Y") +
  theme_minimal()
```


## Introduction



## Summary of Game



## Shot Profile Analysis



## Key Player Performances

The outcome of the Emirates Cup final was shaped largely by the play of each team’s top scorers. Milwaukee’s stars delivered on expectations. Giannis Antetokounmpo led the Bucks with 20 points on 52.6% shooting, while Damian Lillard added 17 points and knocked down 5 of 10 from three. Both players posted efficient shot profiles and outperformed their expected shot quality.

```{r Getting data for star players, echo=FALSE, include=FALSE}
# Manually define your key players (adjust names if needed)
stars <- c("Antetokounmpo, Giannis", "Lillard, Damian",
           "Gilgeous-Alexander, Shai", "Williams, Jalen")

# Filter and prepare table
star_table <- df %>%
  filter(player_nba_shooter %in% stars) %>%
  group_by(player_nba_shooter, team_nba_off) %>%
  summarise(
    FGA = sum(fga, na.rm = TRUE),
    FGM = sum(fg, na.rm = TRUE),
    FG_percent = round(FGM / FGA, 3),
    FG3A = sum(fga3, na.rm = TRUE),
    FG3M = sum(fg3, na.rm = TRUE),
    FG3_percent = round(FG3M / FG3A, 3),
    eFG_percent = round((FGM + 0.5 * FG3M) / FGA, 3),
    Points = sum(fg2 * 2 + fg3 * 3, na.rm = TRUE),
    avg_qSQ = round(mean(qSQ, na.rm = TRUE), 3)
  ) %>%
  ungroup() %>%
  rename(
    Player = player_nba_shooter,
    Team = team_nba_off,
    `FG%` = FG_percent,
    `3PA` = FG3A,
    `3PM` = FG3M,
    `3P%` = FG3_percent,
    `eFG%` = eFG_percent,
    `qSQ` = avg_qSQ
  ) %>%
  arrange(desc(Points))  # or sort however you'd like
```


```{r Star players table, echo=FALSE}
# Display table using gt
star_table %>%
  gt() %>%
  tab_header(
    title = md("**Star Player Comparison**"),
    subtitle = "Shot performance metrics for Bucks and Thunder primary scorers"
  ) %>%
  fmt_percent(columns = c(`FG%`, `3P%`, `eFG%`), decimals = 1) %>%  # ← no qSQ here
  fmt_number(columns = c(`qSQ`), decimals = 3) %>%  # ← format qSQ as a decimal
  fmt_number(columns = c(FGA, FGM, `3PA`, Points), decimals = 0) %>%
  cols_align(align = "center") %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = Player)
  ) %>% 
  tab_source_note(
    source_note = md("*Point totals reflect only field goals made; free throws are not included.*")
  )

```


```{r Creating Player-level shooting summary, include=FALSE}
# Ensure key binary shot variables are numeric
df <- df %>%
  mutate(
    fga = as.numeric(as.character(fga)),
    fg = as.numeric(as.character(fg)),
    fga3 = as.numeric(as.character(fga3)),
    fg3 = as.numeric(as.character(fg3)),
    fg2 = as.numeric(as.character(fg2)),
    fga2 = as.numeric(as.character(fga2))
  )

# Player-level shooting summary
player_summary <- df %>% 
  group_by(player_nba_shooter, team_nba_off) %>% 
  summarise(
    
    FGA = sum(fga, na.rm = TRUE),
    FGM = sum(fg, na.rm = TRUE),
    FG_percent = round(FGM / FGA, 3),
    
    FG2A = sum(fga2, na.rm = TRUE),
    FG2M = sum(fg2, na.rm = TRUE),
    FG2_percent = round(FG2M / FG2A, 3),
    
    FG3A = sum(fga3, na.rm = TRUE),
    FG3M = sum(fg3, na.rm = TRUE),
    FG3_percent = round(FG3M / FG3A, 3),
    
    eFG_percent = round((FGM + 0.5 * FG3M) / FGA, 3),
    avg_qSQ = round(mean(qSQ, na.rm = TRUE), 3),
    
    Points = FG2M * 2 + FG3M * 3
  ) %>%
  arrange(desc(FGA))
```

```{r Setting up players by FGA for display, include=FALSE}
# players by FGA for display
player_summary <- player_summary %>%
  mutate(player_nba_shooter = reorder(player_nba_shooter, FGA))

# Define color by team
bar_colors <- ifelse(player_summary$team_nba_off == "MIL", "#00471B", "#007AC1")
```

```{r Field Goal Attempts by Player bar Chart, echo=FALSE}
# Create highchart
highchart() %>%
  hc_chart(type = "bar") %>%
  hc_title(text = "Field Goal Attempts by Player") %>%
  hc_xAxis(
    categories = player_summary$player_nba_shooter,
    title = list(text = "Player")
  ) %>%
  hc_yAxis(
    title = list(text = "Field Goal Attempts")
  ) %>%
  hc_plotOptions(
    bar = list(dataLabels = list(enabled = TRUE))
  ) %>%
  hc_tooltip(
    pointFormat = paste(
      "<b>Team:</b> {point.custom.team}<br>",
      "<b>FG%:</b> {point.custom.fg_percent}<br>",
      "<b>FG3%:</b> {point.custom.fg3_percent}<br>",
      "<b>eFG%:</b> {point.custom.efg_percent}<br>",
      "<b>Points:</b> {point.custom.points}<br>"
    ),
    useHTML = TRUE
  ) %>%
  hc_add_series(
    data = lapply(1:nrow(player_summary), function(i) {
      list(
        y = player_summary$FGA[i],
        color = bar_colors[i],
        custom = list(
          team = player_summary$team_nba_off[i],
          fg_percent = player_summary$FG_percent[i],
          fg3_percent = player_summary$FG3_percent[i],
          efg_percent = player_summary$eFG_percent[i],
          points = player_summary$Points[i]
        )
      )
    }),
    type = "bar",
    showInLegend = FALSE
  )
```


On the other side, Oklahoma City’s stars struggled to convert. Shai Gilgeous-Alexander and Jalen Williams combined for 44 shot attempts, but shot just 33.3% and 40.0% from the field respectively, totaling 35 points combined. Shai went 2-of-9 from three, while Williams hit only 1-of-4.

Beyond the stars, Milwaukee’s supporting cast provided critical contributions. Brook Lopez and Bobby Portis combined for 20 points on 13 shot attempts, both shooting above 50%. Portis added an efficient interior presence, while Lopez spaced the floor effectively.

For the Thunder, Gary Trent Jr. and Isaiah Joe had opportunities, but were less efficient. They combined for 18 shot attempts and only 5 made field goals. The Thunder’s role players, like the stars, struggled to convert shot quality into scoring output.

Milwaukee’s depth and shooting consistency - not just from their stars, but across the rotation - proved decisive in their Emirates Cup win.


## Five Key Observations



## Conclusion



## Appendix











































```{r}
df %>%
  group_by(team_nba_off) %>%
  summarise(made = sum(fg == 1), missed = sum(fg == 0)) %>%
  hchart("bar", hcaes(x = team_nba_off, y = made), name = "Made Shots")
```






```{r}
interactive_df <- df %>%
  filter(fg3 == 1) %>%
  mutate(
    label = paste0(
      "<b>Shooter:</b> ", player_nba_shooter, "<br>",
      "<b>Distance:</b> ", round(distance, 1), " ft<br>",
      "<b>Contest Level:</b> ", contestLevel, "<br>",
      "<b>Defender Distance:</b> ", round(closestDefDist, 1), " ft"
    )
  )



highchart() %>%
  hc_chart(type = "scatter") %>%
  
  hc_add_series(
    data = interactive_df,
    type = "scatter",
    hcaes(x = loc_x, y = loc_y),
    tooltip = list(pointFormat = "{point.label}"),
    name = "Made Shots"
  ) %>%
  
  hc_title(text = "Interactive Shot Chart: Made Shots Only") %>%
  hc_xAxis(title = list(text = "Court X"), min = -25, max = 25) %>%
  hc_yAxis(title = list(text = "Court Y"), min = 0, max = 47) %>%
  hc_tooltip(useHTML = TRUE) %>%
  hc_plotOptions(scatter = list(marker = list(radius = 4)))
```






























```{r}
# Choose one team to display
selected_team <- "OKC"  # Change this to any team in your dataset

# Filter dataset
team_data <- df %>%
  filter(team_nba_off == selected_team, fg3 == 1) %>%
  mutate(
    tooltip = paste0(
      "<b>Shooter:</b> ", player_nba_shooter, "<br>",
      "<b>Distance:</b> ", round(distance, 1), " ft<br>",
      "<b>Contest:</b> ", contestLevel, "<br>",
      "<b>Defender:</b> ", round(closestDefDist, 1), " ft"
    )
  )

# Load and rasterize court image
court_url <- "https://thedatagame.com.au/wp-content/uploads/2016/03/nba_court.jpg"
court_img <- rasterGrob(readJPEG(getURLContent(court_url)), 
                        width = unit(1, "npc"), height = unit(1, "npc"))

# Create ggplot with interactivity
gg <- ggplot(team_data, aes(x = loc_x, y = loc_y)) +
  annotation_custom(court_img, xmin = -25, xmax = 25, ymin = 0, ymax = 47) +
  geom_point_interactive(aes(tooltip = tooltip), color = "blue", size = 2, alpha = 0.7) +
  coord_fixed(xlim = c(-25, 25), ylim = c(0, 47)) +
  labs(title = paste("Interactive Shot Chart -", selected_team)) +
  theme_void()

# Render as interactive plot
girafe(ggobj = gg)
```





































